### 2013-06-15 doc/pgas13/evaluation.tex
More aggressive cutting to get down to 8 pages.

\section{Evaluation}
To evaluate the impact flat combining has on the global data structures, we ran a series of experiments to test the raw performance of the data structures themselves under different workloads, and then measure their impact on performance of two graph benchmarks.
Experiments were run on a cluster of AMD Interlagos processors. Nodes have 32 2.1-GHz cores in two sockets, 64GB of memory, and 40Gb Mellanox ConnectX-2 InfiniBand network cards, connected via a QLogic InfiniBand switch.

@@

On the mixed workload, the stack is able to do matching locally, allowing it to reduce the amount of communication drastically, greatly improving its performance on the mixed workload. Figure~\ref{fig:stack_stats} corroborates this, showing that the amount of combining that occurs directly correlates with the observed throughput.

We also experimented with performing both forms of combining together; however, combining at the master did not affect performance significantly. This was somewhat surprising as combined updates still force others to wait. However, with both levels of combining enabled, relatively few requests are combined at the master--for the stack, a maximum of 6\% per pass for the all-push workload and 0.1\% for the mixed workload.

@@

% Without combining, performance scales out to 32 nodes because synchronization is distributed over hash cells, but drops off as the number of destinations increases.
Figure~\ref{fig:hash_perf} shows the throughput results for the Set and Map.
Both data structures synchronize at each hash cell, which allows them to scale fairly well even without combining. However, after 32 nodes, scaling drops off significantly due to the increased number of destinations.
Combining allows duplicate inserts and lookups to be eliminated, so performs better the smaller the key range. This reduction in message traffic allows scaling to continue out to 64 nodes.

@@

Figure~\ref{fig:bfs_perf} shows the results of scaling up to 64 nodes. Using the simple queue without flat combining is completely unscalable, but with FC, it tracks the tuned version. This illustrates that providing a safe, synchronized data structure for initially developing algorithms for PGAS is possible without sacrificing scaling.
% , while further optimizations can be applied incrementally.

@@

The results in Figure~\ref{fig:cc_perf} show that none of these three scale well out to 64 nodes. However, performing combining does improve the performance over the uncombined case. The tuned version outperforms the synchronous version because it is able to build up most of the set locally on each core before merging them at the end of the first phase. An implementation that did not provide synchronized semantics could potentially relax consistency in a more general way, but this is left for future work.

>>>>

\section{Evaluation}
To evaluate the impact flat combining has on the global data structures, we ran a series of experiments to test the raw performance of the data structures themselves under different workloads, and then measure their impact on performance of two graph benchmarks.
Experiments were run on a cluster of AMD Interlagos processors. Nodes have 32 2.1-GHz cores in two sockets, 64GB of memory, and 40Gb Mellanox ConnectX-2 InfiniBand network cards connected via a QLogic switch.

@@

On the mixed workload, the stack is able to do matching locally, allowing it to reduce the amount of communication drastically, greatly improving its performance. Figure~\ref{fig:stack_stats} corroborates this, showing that the amount of combining that occurs directly correlates with the observed throughput.

% % left out for space
% We also experimented with performing both forms of combining together. Combining at the master did not affect performance significantly, which is somewhat surprising as all combined requests must still serialize.
% However, with both levels of combining enabled, relatively few requests are combined at the master--for the stack, a maximum of 6\% per pass for the all-push workload and 0.1\% for the mixed workload.

@@

% Without combining, performance scales out to 32 nodes because synchronization is distributed over hash cells, but drops off as the number of destinations increases.
Figure~\ref{fig:hash_perf} shows the throughput results for the Set and Map.
Both data structures synchronize at each hash cell, which allows them to scale fairly well even without combining. However, after 32 nodes, scaling drops off significantly due to the increased number of destinations.
Combining allows duplicate inserts and lookups to be eliminated, so performs better the smaller the key range. This reduction in message traffic allows scaling out to 64 nodes.

@@

Figure~\ref{fig:bfs_perf} shows the scaling results. Using the simple queue without flat combining is completely unscalable, but with FC, it tracks the tuned version. This illustrates that providing a safe, synchronized data structure for initially developing algorithms for PGAS is possible without sacrificing scaling.
% , while further optimizations can be applied incrementally.

@@

The results in Figure~\ref{fig:cc_perf} show that none of these three scale well out to 64 nodes. However, performing combining does improve the performance over the uncombined case. The tuned version outperforms the synchronous version because it is able to build up most of the set locally on each core before merging them at the end of the first phase. An implementation that did not provide synchronized semantics could potentially relax consistency in a more general way.

### 2013-06-15 doc/pgas13/flat-combining.tex
More aggressive cutting to get down to 8 pages.

Simply by delegating work to another core, locality is improved and synchronization is reduced.
% Reduced synchronization comes from delegating work to another core.
To illustrate the issue, consider the shared synchronous stack shown in Figure~\ref{fig:fc_shared_mem}, with pre-allocated storage and a \texttt{top} pointer protected by a lock. Without flat combining, whenever a thread attempts to push something on the stack, it must acquire the lock, put its value into the storage array, bump the top, and then release the lock. When many threads contend for the lock, all but one will fail and have to retry. Each attempt forces an expensive memory fence and consumes bandwidth, and as the number of threads increases, the fraction of successes plummets. Under flat combining, instead, threads add requests to a \emph{publication list}. They each try to acquire the lock, and the one that succeeds becomes the combiner. Instead of retrying, the rest spin on their request waiting for it to be filled. The combiner walks the publication list, performs all of the requests, and when done, releases the lock. This allows the one thread to keep the data structure in cache, reducing thrashing between threads. It also greatly reduces contention on the lock, but introduces a new point of synchronization--adding to the publication list. However, if a thread performs multiple operations, it can leave its publication record in the list and amortize the synchronization cost. This publication list mechanism can be re-used in other data structures, saving each from needing to implement its own clever synchronization.

@@

As mentioned before, the Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine.
% As mentioned before, the local \emph{publication list} is trivial to access atomically in Grappa's cooperative scheduling.
% Instead, in the Grappa FC Framework, 
A \emph{proxy} object is instantiated ahead of time on each core and each worker merges its request into that structure before blocking or becoming the combiner.
Each global data structure must define a \emph{proxy} with the following:
\begin{enumerate}
  \item A local \emph{proxy} data structure for tracking combined requests.
  \item \emph{Combining methods} that operate on the local proxy.
  \item A \emph{sync} method that globally commits the proxy's state.
\end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:proxy}.

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
class GlobalStack {
  GlobalAddress<T> top;
  Grappa::Mutex lock;
};

void push(GlobalAddress<GlobalStack> master, T e){
  // perform operation on the stack's master core
  delegate::call(master.core(), [master,e]{
    // force other operations to wait
    master->lock.acquire();
    // do remote write to the top of the stack
    delegate::write(master->top, e);
    // bump top and release lock
    master->top++;
    master->lock.release();
  }); // blocks until response arrives
}
\end{lstlisting}
\caption{Snippet of code for pushing to a global stack (without flat combining).}
\label{fig:push}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
class GlobalStackProxy : Grappa::FCProxy {
  GlobalAddress<GlobalStack> master;
  // Local state for tracking requests
  T  pushed_values[1024];
  T* popped_results[1024];
  int npush, npop;

@@

  // Global sync (called by FC framework)
  void sync() override {
    if (npush > 0) {
      // on master: acquire lock, return top ptr
      auto top = delegate::call(master.core(),[=]{
        master->lock.acquire();
        return master->top;
      });
      // copy values to top of stack
      Grappa::memcpy(top, pushed_values, npush);
      // bump top and release lock
      delegate::call(master.core(),[master,npush]{
        master->top += npush;
        master->lock.release();
      });
    } else if (npop > 0) {} // elided for space...
  }
};
\end{lstlisting}
\caption{Snippet of the Proxy object for the GlobalStack. The code to synchronize pushes should resemble that from Figure~\ref{fig:push} but for multiple elements.}
\label{fig:proxy}
\end{figure}

With this data-structure-specific proxy object, the FC framework does the bookkeeping to block and wake workers, deliver results, manage when to synchronize, and choose the combiner.
The FC framework is responsible for ensuring that all of the combined operations eventually get committed. There are a number of ways progress could be guaranteed, but one of the simplest is to ensure that as long as there are any outstanding requests, at least one worker is committing a combined operation.
When that combined operation finishes, if there are still outstanding requests that have built up in the meantime, another blocked worker is chosen to do another combined synchronization.

@@

\subsection{Global Stack and Queue}
Figure~\ref{fig:proxy} shows an excerpt from the definition of the proxy object for the Grappa GlobalStack. The proxy tracks pushes locally in the \texttt{pushed\_values} array. When \texttt{pop} is called, if there are pushes in the local array, it immediately takes the value and wakes the last worker to push. Otherwise, it adds a pointer to a location on its stack where the \texttt{sync} operation, will eventually write the result. Because of local matching, when a Stack proxy is synchronized, it will have either all pops or all pushes, which makes the implementation of sync straightforward. Note that the operation to synchronize a batch of pushes looks almost identical to the code to do a single push from Figure~\ref{fig:push}.

The GlobalQueue has nearly the same implementation as the stack, but is unable to match locally, so \texttt{sync} first commits the pushes, then the pops.

@@

Our implementation of the GlobalMap matches that of the Set but of course stores values. This does not change the requirements of the proxy except that the return values of lookups must be a copy of the value rather than just a boolean.

>>>>

Simply by delegating work to another core, locality is improved and synchronization is reduced.
Consider the shared synchronous stack shown in Figure~\ref{fig:fc_shared_mem}, with pre-allocated storage and a \texttt{top} pointer protected by a lock. Without flat combining, whenever a thread attempts to push something on the stack, it must acquire the lock, put its value into the storage array, bump the top, and then release the lock. When many threads contend for the lock, all but one will fail and have to retry. Each attempt forces an expensive memory fence and consumes bandwidth, and as the number of threads increases, the fraction of successes plummets. Under flat combining, instead, threads add requests to a \emph{publication list}. They each try to acquire the lock, and the one that succeeds becomes the combiner. Instead of retrying, the rest spin on their request waiting for it to be filled. The combiner walks the publication list, performs all of the requests, and when done, releases the lock. This allows the one thread to keep the data structure in cache, reducing thrashing between threads. It also greatly reduces contention on the lock, but introduces a new point of synchronization--adding to the publication list. However, if a thread performs multiple operations, it can leave its publication record in the list and amortize the synchronization cost. This publication list mechanism can be re-used in other data structures, saving each from needing to implement its own clever synchronization.

@@

As mentioned before, the Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine.
% As mentioned before, the local \emph{publication list} is trivial to access atomically in Grappa's cooperative scheduling.
% Instead, in the Grappa FC Framework, 
For each global structure instance, a \emph{proxy} object is created on each core and each worker merges its request into that structure before blocking or becoming the combiner.
Each global data structure must define a \emph{proxy} that has \emph{state} to track combined requests, \emph{methods} to add new requests, and a way to \emph{sync} the state globally.
% \begin{enumerate}
%   \item A local \emph{proxy} data structure for tracking combined requests.
%   \item \emph{Combining methods} that operate on the local proxy.
%   \item A \emph{sync} method that globally commits the proxy's state.
% \end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:push}.

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
// Global master state
class GlobalStack {
  GlobalAddress<T> top;
  Grappa::Mutex lock;
};
// Push *without* combining
void push(GlobalAddress<GlobalStack> master, T e){
  // from stack's master core, perform write to top and increment atomically
  delegate::call(master.core(), [master,e]{
    master->lock.acquire();
    delegate::write(master->top, e);
    master->top++;
    master->lock.release();
  }); // blocks until response arrives
}
// Definition of proxy
class GlobalStackProxy : Grappa::FCProxy {
  GlobalAddress<GlobalStack> master;
  // Local state for tracking requests
  T  pushed_values[1024];
  T* popped_results[1024];
  int npush, npop;

@@

  // Global sync (called by FC framework)
  void sync() override {
    if (npush > 0) {
      // on master: acquire lock, return top ptr
      auto top = delegate::call(master.core(),[=]{
        master->lock.acquire();
        return master->top;
      });
      // copy values to top of stack
      Grappa::memcpy(top, pushed_values, npush);
      // bump top and release lock
      delegate::call(master.core(),[master,npush]{
        master->top += npush;
        master->lock.release();
      });
    } else if (npop > 0) {} // elided for space...
  }
};
\end{lstlisting}
\caption{Snippet of code from GlobalStack.
% Note how the combined \texttt{push} resembles the uncombined version but for multiple elements.
}
\label{fig:push}
\end{figure}

% The FC framework does the bookkeeping to block and wake workers, deliver results, manage when to synchronize, and choose the combiner.
The FC framework is responsible for ensuring that all of the combined operations eventually get committed. There are a number of ways progress could be guaranteed, but one of the simplest is to ensure that as long as there are any outstanding requests, at least one worker is committing a combined operation.
When that combined operation finishes, if there are still outstanding requests that have built up in the meantime, another blocked worker is chosen to do another combined synchronization.

@@

\subsection{Global Stack and Queue}
Figure~\ref{fig:push} shows an excerpt from the definition of the proxy object for the Grappa GlobalStack. The proxy tracks pushes in the \texttt{pushed\_values} array. When \texttt{pop} is called, if there are pushes, it immediately takes the top value and wakes the last pusher. Otherwise, it adds a pointer to a location on its stack where the \texttt{sync} operation will write the result. Because of local matching, when a Stack proxy is synchronized, it will have either all pops or all pushes, which makes the implementation of sync straightforward. Note that the operation to synchronize a batch of pushes looks almost identical to the code to do a single push from Figure~\ref{fig:push}.

The GlobalQueue has nearly the same implementation as the stack, but is unable to match locally.
%, so \texttt{sync} first commits the pushes, then the pops.

@@

Our implementation of the GlobalMap matches that of the Set but of course stores values.
 % This does not change the requirements of the proxy except that the return values of lookups must be a copy of the value rather than just a boolean.

### 2013-06-15 doc/pgas13/intro.tex
More aggressive cutting to get down to 8 pages.

\section{Introduction}
The goal of partitioned global address space (PGAS)~\cite{upc:2005} languages and runtimes is to provide the illusion of a single shared memory to a program actually executing on a distributed memory machine such as a cluster. This allows programmers to write their algorithms without needing to explicitly manage communication. 
However, with this shared memory abstraction come all of the difficult concurrency issues that arise in physically shared memory. Luckily, the PGAS community can leverage a large body of work solving these issues in shared memory, where the costs are different, providing an opportunity for different trade-offs to be made.

Globally shared data structures are one of the cornerstones of shared-memory and PGAS abstractions. 
In order for multiple concurrent threads to interact with one another correctly, they must observe shared data consistently. It is commonly accepted that the easiest consistency model to reason about is \emph{sequential consistency} (SC), which enforces that all accesses appear to happen in some global serializable order.
To preserve SC, operations on shared data structures should be \emph{linearizable}~\cite{herlihy1990linearizability}; that is, appear to happen atomically in some global total order.
In both physically shared memory and PGAS, maintaining linearizability presents performance challenges.
The simplest way is to have a single global lock for implementing mutual exclusion for all operations, enforcing atomicity and linearizability. The cost of literally serializing accesses in this way is typically considered prohibitive, even in physically shared memory.
With the massive amount of parallelism in a cluster of multiprocessors and with the increased cost of remote synchronization, the problem is magnified.

Much prior work has explored the scalability challenges of adding more concurrent accessors to shared data structures in shared memory systems.
In the classic multithreading case, increasing concurrency increases contention and damages performance.
Either all threads fight to obtain a single global lock, or in more sophisticated lock-free schemes, they attempt a series of contentious atomic operations, and retry if they fail. In both cases, as the number of concurrent accessors increases, the more they contend and the more synchronization operations fail.

A new synchronization technique called \emph{flat combining}~\cite{flatCombining} coerces threads to \emph{cooperate} rather than \emph{contend}.
Threads delegate their work to a single thread, giving it the opportunity to combine multiple requests in data-structure specific ways and perform them free from contention.
Together, the reduced synchronization cost and combining allow even a data structure with a single global lock to scale better than complicated concurrent data structures using fine-grained locking or lock-free mechanisms.

The goal of this work is to apply the flat-combining paradigm to a PGAS runtime to reduce the cost of maintaining sequentially consistent data structures.
We leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
By leveraging lots of concurrency, we find that it is actually possible to scalably provide sequential consistency. We leverage Grappa's latency tolerance mechanisms to allow many fine-grained synchronized operations to be combined to achieve higher overall throughput.
In addition, we show how a generic flat-combining framework can be used to implement multiple global data structures.

The next section will describe in more detail the Grappa runtime system that is used to implement flat combining for distributed memory machines. We then explain the flat-combining paradigm in more depth and describe how it maps to a PGAS model. Next, we explain how several data structures are implemented in our framework and show how they perform on simple throughput workloads. Finally, we evaluate how this affects application performance in two simple graph analysis kernels.

>>>>

\section{Introduction}
The goal of partitioned global address space (PGAS)~\cite{upc:2005} languages and runtimes is to provide the illusion of a single shared memory to a program actually executing on a distributed memory machine such as a cluster. This allows programmers to write their algorithms without needing to explicitly manage communication.
However, it does not alleviate the need for reasoning about consistency among concurrent threads.
Luckily, the PGAS community can leverage a large body of work solving these issues in shared memory and explore how differing costs lead to different design choices.

% Globally shared data structures are one of the cornerstones of shared-memory and PGAS abstractions. 
% In order for multiple concurrent threads to interact with one another correctly, they must observe shared data consistently.
It is commonly accepted that the easiest consistency model to reason about is \emph{sequential consistency} (SC), which enforces that all accesses appear to happen in some global serializable order.
To preserve SC, operations on shared data structures should be \emph{linearizable}~\cite{herlihy1990linearizability}; that is, appear to happen atomically in some global total order.
In both physically shared memory and PGAS, maintaining linearizability presents performance challenges.
The simplest way is to have a single global lock for implementing mutual exclusion for all operations, enforcing atomicity and linearizability. The cost of literally serializing accesses in this way is typically considered prohibitive, even in physically shared memory.
% Much prior work has explored the scalability challenges of adding more concurrent accessors to shared data structures in shared memory systems.
In the classic multithreading case, as the number of concurrent accessors increases, the more they contend and the more synchronization operations fail.
With the massive amount of parallelism in a cluster of multiprocessors and with the increased cost of remote synchronization, the problem is magnified.

A new synchronization technique called \emph{flat combining}~\cite{flatCombining} coerces threads to \emph{cooperate} rather than \emph{contend}.
Threads delegate their work to a single thread, giving it the opportunity to combine multiple requests in data-structure specific ways and perform them free from contention.
This allows even a data structure with a single global lock to scale better than complicated concurrent data structures using fine-grained locking or lock-free mechanisms.

The goal of this work is to apply the flat-combining paradigm to a PGAS runtime to reduce the cost of maintaining sequentially consistent data structures.
We leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
We leverage Grappa's latency tolerance mechanisms to allow many fine-grained synchronized operations to be combined to achieve higher, scalable throughput while maintaining sequential consistency.
In addition, we show how a generic flat-combining framework can be used to implement multiple global data structures.

The next section will describe in more detail the Grappa runtime system that is used to implement flat combining for distributed memory machines. We then explain the flat-combining paradigm in more depth and describe how it maps to a PGAS model. Next, we explain how several data structures are implemented in our framework and show how they perform on simple throughput workloads as well as in two simple graph analysis kernels.

### 2013-06-15 doc/pgas13/relatedwork.tex
More aggressive cutting to get down to 8 pages.

\section{Related Work}
Attempts to improve the scalability of synchronized data structures are as old as the data structures themselves. Though much attention has gone to lock-free data structures such as the Treiber Stack~\cite{treiber} and Michael-Scott Queue~\cite{msqueue}, a significant body of work has explored ways of scaling globally locked data structures.

Many techniques performed combining hierarchically to reduce synchronization and hot-spotting, differing mainly in the structures used to do combining: fixed trees~\cite{yew:combining-trees}, dynamic trees (or ``funnels'')~\cite{funnels,MAMA}, and randomized trees~\cite{edtrees}. In particular, MAMA~\cite{MAMA}, built for the MTA-2, leveraged funnels to better utilize hardware optimized for massive concurrency.
On the other hand, \emph{flat} combining~\cite{flatCombining} observed that in multicore systems, hierarchical schemes have too much overhead, and a cleverly implemented \emph{flat} publication list can perform better.
Follow-on work introduced parallel flat-combining~\cite{scalableFCQueues}, in which multiple publication lists are combined in parallel and their remainders serialized.
Flat combining was also applied in the NUMA (non-uniform memory access) domain with scalable hierarchical locks~\cite{fcNUMALocks}, which improves on prior work on hierarchical locking by leveraging flat-combining's publication mechanism which amortizes the cost of synchronization.

Our work extends the flat-combining paradigm further, to the PGAS domain, where only software provides the illusion of a physically distributed memory. The relatively higher cost of communication in PGAS makes flat combining even more compelling. Within-core combining is naturally done in a \emph{flat} way. In addition, the physical isolation per node combined with Grappa's engineered per-core isolation guarantees, create a novel situation for combining, which led to our \emph{combining proxy} design.

In the distributed memory and PGAS domains, implementing scalable and distributed synchronization is the name of the game.
Besides Grappa, other work has proposed leveraging latency tolerance in PGAS runtimes, including MADNESS~\cite{shet:async-upc} which proposes ways of using asynchronous operations in UPC, and Zhang et al.~\cite{zhang:barnes-hut} who use asynchronous messages in a UPC Barnes-Hut implementation to overlap computation with communication and allow for buffering.
Jose et al. introduced UPC Queues~\cite{jose:upc-queues}, a library which provides explicit queues as a replacement for locks to synchronize more efficiently and allow for buffering communication. This work demonstrates a similar use case for synchronized queues as in our work, and uses them with the Graph500 benchmark as well. In contrast, Grappa performs much of the same message aggregation invisibly to the programmer, and the FC framework can be used to implement many different linearizable data structures.

>>>>

\section{Related Work}
% Attempts to improve the scalability of synchronized data structures are as old as the data structures themselves.
Though much attention has gone to lock-free data structures such as the Treiber Stack~\cite{treiber} and Michael-Scott Queue~\cite{msqueue}, a significant body of work has explored ways of scaling globally locked data structures using combining to reduce synchronization and hot-spotting.
Techniques differ mainly in the structures used to do combining: fixed trees~\cite{yew:combining-trees}, dynamic trees (or ``funnels'')~\cite{funnels,MAMA}, and randomized trees~\cite{edtrees}. In particular, MAMA~\cite{MAMA}, built for the MTA-2, leveraged funnels to better utilize hardware optimized for massive concurrency.
On the other hand, \emph{flat} combining~\cite{flatCombining} observed that in multicore systems, hierarchical schemes have too much overhead, and a cleverly implemented \emph{flat} publication list can perform better.
Follow-on work introduced parallel flat-combining~\cite{scalableFCQueues}, in which multiple publication lists are combined in parallel and their remainders serialized.
Flat combining was also applied in the NUMA (non-uniform memory access) domain with scalable hierarchical locks~\cite{fcNUMALocks}, which improves on prior work on hierarchical locking by leveraging flat-combining's publication mechanism which amortizes the cost of synchronization.

Our work extends the flat-combining paradigm further, to the PGAS domain, where only software provides the illusion of a physically distributed memory. The relatively higher cost of communication in PGAS makes flat combining even more compelling. In addition, the physical isolation per node combined with Grappa's engineered per-core isolation guarantees, create a novel situation for combining, which led to our \emph{combining proxy} design.

In the distributed memory and PGAS domains, implementing scalable and distributed synchronization is the name of the game.
Besides Grappa, other work has proposed leveraging latency tolerance in PGAS runtimes, including MADNESS~\cite{shet:async-upc} which proposes ways of using asynchronous operations in UPC, and Zhang et al.~\cite{zhang:barnes-hut} who use asynchronous messages in a UPC Barnes-Hut implementation to overlap computation with communication and allow for buffering.
Jose et al. introduced UPC Queues~\cite{jose:upc-queues}, a library which provides explicit queues as a replacement for locks to synchronize more efficiently and allow for buffering communication. This work demonstrates a similar use case for synchronized queues as in our work, and uses them with the Graph500 benchmark as well. In contrast, Grappa performs much of the same message aggregation invisibly to the programmer, and the FC framework can be used to implement many data structures.
